\mytitile\\~

### 2019-10-11 document.tex

\usepackage{graphicx}% Include figure files
%\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{fixme}
%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{mathptmx}
\usepackage{lipsum}
\usepackage{amsmath}
\usepackage{physics}
\usepackage{xparse}
\usepackage{bbm}
\usepackage{hyperref}
%\usepackage{multirow}
%\usepackage{makecell}
\graphicspath{{Pictures/}}

@@

\newcommand{\mytitile}{Effects of strong interaction between light and matter in a superconducting artificial molecule}
\begin{document}
	\preprint{AIP/123-QED}

@@

	\maketitle
\section{Introduction}
Over the past twenty years, superconducting artificial atoms (SAA) were used in numerous experiments confidently demonstrating the validity of fundamental quantum mechanical laws\cite{you2011atomic}. Due to synthetic nature of SAA, their Hamiltonians can be pre-designed and engineered, and this makes them a particularly versatile tool for studies in quantum optics. Notably, their parameters may be varied in such a wide range that novel physical effects previously inaccessible for natural systems may be directly observed. For example, the strong coupling regime when the coupling strength between light and a quantum system approaches 1\% of their frequency has been a major milestone for atoms, molecules, etc and was only reached using high-finesse cavities. In contrast, superconducting quantum systems have rapidly attained this landmark using \textit{circuit QED}\cite{wallraff2004strong} just in a few years after their first discovery\cite{chiorescu2004coherent} and have already surpassed other strongly-coupled systems in terms of coherence\cite{forn2019ultrastrong}. 
Moreover, SAA do not even require confined radiation to implement strong coupling with light: they may be coupled unprecedently strong with the free-propagating EM-waves in the on-chip 1D waveguides\cite{astafiev2010resonance} without having to use cavities at all. The Rabi frequency in this case may reach 50\% of the transition frequency\cite{deng2015observation} which is already the ultra-strong coupling regime.

@@

@@

\subsection{\label{sec:level1} Experimental and numerical parts}
Two tone spectroscopy was found experimentally and shows good agreement with modelling data \autoref{fig:two-tone}.

\begin{figure*}
	\centering
	\includegraphics[width=\linewidth]{main_picture}
	\caption{Two-tone spectroscopy: \textbf{(a)} Experiment;\textbf{(b)} Modelling ; the colorbar is common.  Clear flux-dependent transmon transitions are visible.}
	\label{fig:two-tone}
\end{figure*}

@@

From the simulation follows that a two-level approximation for first qubit and three-level approximation for second is sufficient to observe the effect we are interested in. The Hamiltonian in matrix form is given by the following expression:
\begin{widetext}
	\begin{equation}\label{matrix} 
	\text{$H^R$} = \left(\begin{matrix}
	0 & \text{$\Omega_1/2 $} & \text{$\Omega _2$} e^{i \text{$\delta_{\omega}$} t}/2 & 0 & 0 & 0 \\
	\text{$\Omega_1/2 $} & \text{$-\Delta_1 $} & J & \text{$\Omega_2$} e^{i \text{$\delta_{\omega}$} t}/2 & 0 & 0 \\
	\text{$\Omega_2$} e^{-i \text{$\delta_{\omega} $} t}/2 & J & -\text{$\Delta_2 $} &
	\text{$\Omega_1$/2} & \sqrt{2} \text{$\Omega_2$} e^{i \text{$\delta_{\omega}$} t}/2 & 0 \\
	0 & \text{$\Omega_2$} e^{-i \text{$\delta_{\omega} $} t}/2 & \text{$\Omega_1/2$} &
	\text{$-\Delta_2 -\Delta_1$} & \sqrt{2} J & \sqrt{2} \text{$\Omega_2$} e^{i \text{$\delta_{\omega}
			$} t}/2 \\
	0 & 0 & \sqrt{2} \text{$\Omega_2$} e^{-i \text{$\delta_{\omega}$} t}/2 & \sqrt{2} J &
	\text{$\alpha$}-2 \text{$\Delta_2$} & \text{$\Omega_1/2$} \\
	0 & 0 & 0 & \sqrt{2} \text{$\Omega_2$} e^{-i \text{$\delta_{\omega} $} t}/2 & \text{$\Omega_1/2
		$} & \text{$\alpha$}-2 \text{$\Delta_2$}- \text{$\Delta_1$} \\
	\end{matrix}\right)
	\end{equation}
\end{widetext}
Where $\Omega_1$, $\Omega_2$ - amplitudes of the drives, $\Delta_1 = \omega_d-\omega_{tr1}=0$,$\Delta_2 = \omega_d-\omega_{tr2}$.
Let's consider the Hamiltonian (\autoref{matrix}) in details. Divide it into three parts: time-dependent $(V_t)$, the part responsible for the interaction between the qubits $(V_J)$ and all the rest $(H_0)$.
\begin{equation}
	H^R=H_0+V_J+V_t
\end{equation}
The $\Omega_1\gg \Omega_2,\Delta_2\gg J$, so the driven field on the first qubit is strong and form the dressed states defined by the Hamiltonian $H_0$.
The Hamiltonian $H_0$ is station and we can find eigenvectors and eigenvalues. The drive on the second qubit is weak with very low intensity, so we can neglect any perturbation introduced by second drive on the position, width os the dressed states.
Firstly, we consider the system of non-interacting qubits $(J=0)$. In this case some transitions are prohibited and the splitting we are interested in is not observed. 
Therefore a non-zero interaction between qubits is necessary to observe AT effect.

@@

\section{Theoretical description of two-photon process}

@@

In order to determine which transitions correspond to the lines, the stationary Schrödinger equation was solved \autoref{fig:spectr} in three-level approximation for both of qubits. With the Hamiltonian of the system:
\begin{equation}
	H_{system} = H_{tr1}\otimes\mathbbm{1}+\mathbbm{1}\otimes H_{tr2}+H_{int}
\end{equation}
\begin{figure}[h]
	\centering
	\includegraphics[width=\linewidth]{spectr}
	\caption{Two-tone spectroscopy with transitions lines. In the red frame, four transitions intersect, three of which originate from the state $\ket{00}$ to $\ket{10}$, two-photon
		$\ket{02}$, three-photon
		$\ket{12}$ and the fourth from  $\ket{10}$ to  $\ket{02}$, where $\ket{ij}$ denote the $ij-th$ energy eigenstate of the transmon system, i,j represent first and second qubit respectively}
	\label{fig:spectr}
\end{figure}

>>>>

\usepackage{graphicx}% Include figure files
%\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{fixme}
%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines
\usepackage{hyperref}
\usepackage{kbordermatrix}% http://www.hss.caltech.edu/~kcb/TeX/kbordermatrix.sty
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{mathptmx}
\usepackage{lipsum}
\usepackage{amsmath}
\usepackage{physics}
\usepackage{xparse}
\usepackage{bbm}
\usepackage{nicematrix}
%\usepackage{multirow}
%\usepackage{makecell}
\graphicspath{{Pictures/}}

@@

\newcommand{\mytitile}{Effects of strong interaction between light and matter in a superconducting artificial molecule}

\begin{document}
	\preprint{AIP/123-QED}

@@

	\maketitle
\section{Introduction}

Over the past twenty years, superconducting artificial atoms (SAA) were used in numerous experiments confidently demonstrating the validity of fundamental quantum mechanical laws\cite{you2011atomic}. Due to synthetic nature of SAA, their Hamiltonians can be pre-designed and engineered, and this makes them a particularly versatile tool for studies in quantum optics. Notably, their parameters may be varied in such a wide range that novel physical effects previously inaccessible for natural systems may be directly observed. For example, the strong coupling regime when the coupling strength between light and a quantum system approaches 1\% of their frequency has been a major milestone for atoms, molecules, etc and was only reached using high-finesse cavities. In contrast, superconducting quantum systems have rapidly attained this landmark using \textit{circuit QED}\cite{wallraff2004strong} just in a few years after their first discovery\cite{chiorescu2004coherent} and have already surpassed other strongly-coupled systems in terms of coherence\cite{forn2019ultrastrong}. 
Moreover, SAA do not even require confined radiation to implement strong coupling with light: they may be coupled unprecedently strong with the free-propagating EM-waves in the on-chip 1D waveguides\cite{astafiev2010resonance} without having to use cavities at all. The Rabi frequency in this case may reach 50\% of the transition frequency\cite{deng2015observation} which is already the ultra-strong coupling regime.

@@

The implementations of the transitions between the eigenstates is achieved by external microwave driving through a capacitively coupled transition line. In the notations used above, the pump Hamiltonian can be written generically as $ 	H_{dr} = -2e n\beta V_{dr}(t)=\Omega(t)(b+b^{\dagger})$.
Where $2en$ -- the charge on island, $\beta$ -- dimensionless coefficient depending on the geometry of the system and the ratio of capacities, $V_{dr}(t)$ -- the voltage on the external source; $n =\frac{1}{\sqrt{2}} (\frac{E_J}{8E_C})^{1/4}(b+b^{\dagger})$.

@@

\subsection{\label{sec:level1} Experimental and numerical parts}

Two tone spectroscopy was found experimentally and shows good agreement with modelling data \autoref{fig:two-tone}.

\begin{figure*}

	\centering
	\includegraphics[width=\linewidth]{main_picture}
	\caption{Two-tone spectroscopy: (a) Experiment;b(b) Modelling ; the colorbar is common.  Clear flux-dependent transmon transitions are visible.}
	\label{fig:two-tone}
\end{figure*}

@@

From the simulation follows that a two-level approximation for first qubit and three-level approximation for second is sufficient to observe the effect we are interested in. The Hamiltonian in matrix form is given by the following expression:

\renewcommand{\kbldelim}{(}% Left delimiter
\renewcommand{\kbrdelim}{)}% Right delimiter
\begin{widetext}
	\begin{equation}\label{matrix} 
	\text{$H^R$} = \kbordermatrix{
	\ket{00} & \ket{10} & \ket{01} & \ket{11} & \ket{02} & \ket{12} \\
	0 & \text{$\Omega_1/2 $} & \text{$\Omega _2$} e^{i \text{$\delta_{\omega}$} t}/2 & 0 & 0 & 0 \\
	\text{$\Omega_1/2 $} & \text{$-\Delta_1 $} & J & \text{$\Omega_2$} e^{i \text{$\delta_{\omega}$} t}/2 & 0 & 0 \\
	\text{$\Omega_2$} e^{-i \text{$\delta_{\omega} $} t}/2 & J & -\text{$\Delta_2 $} &
	\text{$\Omega_1$/2} & \sqrt{2} \text{$\Omega_2$} e^{i \text{$\delta_{\omega}$} t}/2 & 0 \\
	0 & \text{$\Omega_2$} e^{-i \text{$\delta_{\omega} $} t}/2 & \text{$\Omega_1/2$} &
	\text{$-\Delta_2 -\Delta_1$} & \sqrt{2} J & \sqrt{2} \text{$\Omega_2$} e^{i \text{$\delta_{\omega}
			$} t}/2 \\
	0 & 0 & \sqrt{2} \text{$\Omega_2$} e^{-i \text{$\delta_{\omega}$} t}/2 & \sqrt{2} J &
	\text{$\alpha$}-2 \text{$\Delta_2$} & \text{$\Omega_1/2$} \\
	0 & 0 & 0 & \sqrt{2} \text{$\Omega_2$} e^{-i \text{$\delta_{\omega} $} t}/2 & \text{$\Omega_1/2
		$} & \text{$\alpha$}-2 \text{$\Delta_2$}- \text{$\Delta_1$} \\
}
	\end{equation}
\end{widetext}
Where $\Omega_1$, $\Omega_2$ - amplitudes of the drives, $\Delta_1 = \omega_d-\omega_{tr1}=0$,$\Delta_2 = \omega_d-\omega_{tr2}$.
Let's consider the Hamiltonian (\autoref{matrix}) in details. Divide it into three parts: time-dependent $(V_t)$, the part responsible for the interaction between the qubits $(V_J)$ and all the rest $(H_0)$.
\begin{equation}
	H^R=H_0+V_J+V_t
\end{equation}
The $\Omega_1\gg \Omega_2,\Delta_2\gg J$, so the driven field on the first qubit is strong and form the dressed states defined by the Hamiltonian $H_0$.
The Hamiltonian $H_0$ is station and we can find eigenvectors and eigenvalues. The drive on the second qubit is weak with very low intensity, so we can neglect any perturbation introduced by second drive on the position, width os the dressed states.
Firstly, we consider the system of non-interacting qubits $(J=0)$. In this case some transitions are prohibited and the splitting we are interested in is not observed. 
Therefore a non-zero interaction between qubits is necessary to observe AT effect.

@@

\section{Theoretical description of two-photon process}
Perturbation theory allows semianalytical calculation of the transition rates (Rabi oscillation frequencies) for single and multi-photon processes.

Let's consider the Hamiltonian $H = H_0+V(t)$, with time-dependent interaction $V(t)$ and "reference" part -- $H_0$ and wave function $\Psi(t)$ . After switching to the interaction picture $V(t)$ modifies to $\tilde{V}(t) = \exp(i H_0 t)V(t)\exp(-i H_0 t)$ and Schrödinger equation gives by $i \derivative{\psi(t)}{t}=\tilde{V}(t)\psi(t)$, where $\psi(t)=\exp(i H_0 t)\Psi(t)$. The reference states of $H_0$ will be denoted $\ket{j}$, $H_0 \ket{j} = \omega_j \ket{j}$ whence it follows that matrix elements  of the modified interaction take the form $\bra{j}\tilde{V}(t)\ket{i}=\exp(i(\omega_j-\omega_i)t)\bra{j}V(t)\ket{i}$. A "solution" of Schrödinger equation in interaction picture corresponding to the initial reference state $\ket{i}$ is
\begin{equation}\label{forpsi}
	\psi_i(t)=\ket{i} - i \int_{-\infty}^{t}dt_1\tilde{V}(t_1)\ket{\psi_i(t_1)}
\end{equation}

The quantity of primary interest is not the wave function but the probability of transition from the initial state $\ket{i}\rightarrow\ket{f}$, where $\ket{f}$ is the final state of interest. Firstly, we find the transition amplitude $A_{fi} = \bra{f}U(t,t_0)\ket{i}=\bra{f}\psi_i(t)$ where $U(t,t_0)$ evolution operator.

Solving \autoref{forpsi} by simple iteration gives the series solution
\begin{equation}
	U(t,t_0) = 1+\sum_{n}U^{(n)}(t,t_0)
\end{equation}
\begin{equation}
	U^{(n)}(t,t_0) = -i^n \int_{t_0}^{t}d t_1\tilde{V}(t_1)...\int_{t_0}^{t_{n-1}}d t_n\tilde{V}(t_n)
\end{equation}
For "weak" interaction the series may be truncated to a fine term $n$ and the $n$th- order transition amplitude $\bra{f}U^{(n)}(t,t_0)\ket{i}$ can be evaluated. 
In our case the interaction in interaction picture satisfies the following relation
\begin{equation}
	\begin{split}
	\tilde{V}(t)=\exp(i H_0 t) \Omega\cos(\omega_d t)(\hat b+\hat b^{\dagger})\exp(-i H_0 t)\overset{RWA}{=}\\ \exp(i H_0 t) \Omega(\hat b^{\dagger} e^{-i\omega_d t}+\hat be^{i\omega_d t})\exp(-i H_0 t)
	\end{split}
\end{equation}
Taking the matrix element between the initial and the final states $\ket{i}$ and $\ket{f}$, one immediately obtains
\begin{equation}
	\begin{split}
	\bra{f}U^{(1)}(t,t_0)\ket{i}=-i\int_{t_0}^{t}dt_1\tilde{V}(t_1)=\\-i\frac{\Omega}{2}\big[\int_{t_0}^{t}dt_1e^{i(\omega_f-\omega_i-\omega_d)t_1}\bra{f}b^{\dagger}\ket{i}+\\
	\int_{t_0}^{t}dt_1e^{i(\omega_f-\omega_i+\omega_d)t_1}\bra{f}b\ket{i}\big]
	\end{split}
\end{equation}
To calculate second-order term we used the following expression $\sum_{j}\ket{j}\bra{j}=\mathbbm{1}$.
\begin{equation}
	\begin{split}
	\bra{f}U^{(2)}(t,t_0)\ket{i}=-\bra{f}\int_{t_0}^{t}dt_1\tilde{V}(t_1)\mathbbm{1}\int_{t_0}^{t_1}dt_2\tilde{V}(t_2)\ket{i}=\\-\int_{t_0}^{t}dt_1 e^{i(\omega_f-\omega_j)t_1}\bra{f}V(t_1)\ket{j}\int_{t_0}^{t_1}dt_2e^{i(\omega_j-\omega_i)t_2}\bra{j}V(t_2)\ket{i}
	\end{split}
\end{equation}
Go over the long-time limit $t\rightarrow\infty$ in the final integral and collect the terms belongings to the same delta functions. This gives
\begin{equation}
	\begin{split}
	\bra{f}U^{(2)}(\infty)\ket{i}=\\\frac{\pi i\Omega^2}{2}\bigg[\delta(\omega_f-\omega_i+2\omega_d)\sum_j\frac{\bra{f}b\ket{j}\bra{j}b\ket{i}}{\omega_j-\omega_i+\omega_d}+\\
	\delta(\omega_f-\omega_i)\bigg(\sum_j\frac{\bra{f}b^{\dagger}\ket{j}\bra{j}b\ket{i}}{\omega_j-\omega_i+\omega_d}+\\
	\sum_j\frac{\bra{f}b\ket{j}\bra{j}b^{\dagger}\ket{i}}{\omega_j-\omega_i-\omega_d}\bigg)+\\
	\delta(\omega_f-\omega_i-2\omega_d)\sum_j\frac{\bra{f}b^{\dagger}\ket{j}\bra{j}b^{\dagger}\ket{i}}{\omega_j-\omega_i-\omega_d}\bigg]
	\end{split}
\end{equation}

@@

In order to determine which transitions correspond to the lines, the stationary Schrödinger equation was solved \autoref{fig:spectr} in three-level approximation for both of qubits. With the Hamiltonian of the system:
\begin{equation}
	H_{system} = H_{tr1}\otimes \mathbbm{1}+\mathbbm{1}\otimes H_{tr2}+H_{int}
\end{equation}
\begin{figure}[h]
	\centering
	\includegraphics[width=\linewidth]{spectr}
	\caption{Two-tone spectroscopy with transitions lines. In the red frame, four transitions intersect, three of which originate from the state $\ket{00}$ to $\ket{10}$, two-photon
		$\ket{02}$, three-photon
		$\ket{12}$ and the fourth from  $\ket{10}$ to  $\ket{02}$, where $\ket{ij}$ denote the $ij-th$ energy eigenstate of the transmon system, i,j represent first and second qubit respectively}
	\label{fig:spectr}
\end{figure}
